---
title: "Project proposal"
author: "Ella and Leo"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(readr)

```

## 1. Introduction

For our project we are working with a data set called "Youth Risk Behavior Surveillance System," collected by the CDC. This data set looks at demographics, environments, and behaviors in teenagers from 9-12th grade across the country. According to the CDC the intent of this data is to "determine how often unhealthy behaviors occur,assess whether unhealthy behaviors increase, decrease, or stay the same over time,provide data at the national, state, territorial and freely associated state, tribal, and local levels, and provide data comparing different groups of adolescents. The data is collected through surveys given to students at both public and private high schools across the US. 

Due to the size of the data we are looking at data from Maine only in the years 2015 and 2017. The behaviors (variables) we will be looking at specifically are:

 [1] "Behaviors that Contribute to Unintentional Injuries"
 [2] "Behaviors that Contribute to Violence"              
 [4] "Cigarette Use"                                      
 [5] "Other Tobacco Use"                                  
 [6] "Alcohol Use"                                        
 [7] "Other Drug Use"   
 [8] "Sexual Behaviors"    
 [9] "Physical Activity"                                  


There are several questions we are interested in exploring through this data:

1. What is the relationship between tobacco use and physical activity? 
2. Are certain demographics of students more likely to engage in risky behaviors?
3. How does age of initiation of use of a specific substance correlate with current use of that substance?
4. Are students who use tobacco more likely to use alcohol or marijuana?



## 2. Data

```{r load-data}
youth_risk <- read_csv("../data/DASH_-_Youth_Risk_Behavior_Surveillance_System__YRBSS___High_School____Including_Sexual_Orientation_20240207.csv")
# View(youth_risk)

```

```{r}
glimpse(new_youth_risk_df)
```

```{r}
new_youth_risk_df <- youth_risk %>% 
  select("YEAR", "Topic", "Subtopic", "ShortQuestionText", "Greater_Risk_Data_Value", "Lesser_Risk_Data_Value", "Sample_Size", "Sex", "Race", "Grade", "SexualIdentity", "SexOfSexualContacts")

```

```{r}
glimpse(new_youth_risk_df)
```

```{r}
# new_youth_risk_df %>%
#   filter(Subtopic != c("Suicide-Related Behaviors", "Fruit and fruit juices", "Milk", "Obesity and Overweight", "Sleep")) 
```


```{r}
new_youth_risk_df %>%
  filter(Subtopic == "Milk") %>%
  group_by(ShortQuestionText) %>%
  mutate(Greater_Risk_Data_Value = round(Greater_Risk_Data_Value, digits = 0)) %>%
  arrange(desc(Greater_Risk_Data_Value))
  
```

```{r}
new_youth_risk_df %>%
  filter(Subtopic == "Milk") %>%
  group_by(ShortQuestionText, Race) %>%
  summarise(Greater_Risk_Data_Value = round(mean(Greater_Risk_Data_Value, na.rm =TRUE), digits = 0)) %>%
  arrange(desc(Greater_Risk_Data_Value)) %>%
  filter(Race != "Total")
```


## 3. Ethics review



## 4. Data analysis plan


We would first like to look at distribution in data biased on gender, race, and grade (see how many males vs. females, etc.). This visualization will probably be a scatter plot and will tell us if the demographics represented in the data seem proportional.
Then we will look at the correlation between tobacco use and physical activity and asthma. We will most likely visualize this in a density plots or a barplot.
Then we will look at tobacco use vs drug/alcohol use. We will most likely visualize this in a barplot.
Then we will look at the correlation between the initiation of alcohol use compared to current alcohol use + demographics (gender, race, and grade). We will most likely visualize this in a density plots or a barplot.
Then we will look at the correlation between the initiation of marijuana use vs current marijuana use + demographics (gender, race, and grade). We will most likely visualize this in a density plots or a barplot.
>>>>>>> d75cc983c464e3381b77908b303770c2203166d1
=======


